The Importance of Initial Conditions in Seasonal Predictions of Antarctic Sea Ice

### 2025-12-11 analysis/paper/access-ice.qmd
Adds section title to competing interests and etc...

bibliography: references.bib  
running:
  title: The Importance of Initial Conditions in Seasonal Predictions of Antarctic Sea Ice
  author: Campitelli et al.
# This section is mandatory even if you declare that no competing interests are present.
competinginterests: |
  The authors declare no competing interests.
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
copyrightstatement: |
  The author's copyright for this publication is transferred to institution/company. 
### The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
### It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.
### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
availability:
  #code: |
  #  use this to add a statement when having only software code available
  #data: |
  #  use this to add a statement when having only data sets available
  codedata: |
    The underlying code for this study is available on GitHub: https://github.com/eliocamp/access-s2_ice-eval.
    Raw data of +S1 and +S2 hindcast are not available due to size. 
    Derived datasets required to reproduce the results, including extent timeseries and error measures, are available in this Zenodo repository: https://zenodo.org/records/17479538 [@campitelli2025]
authorcontribution: |
    EC performed the data analysis and wrote the manuscript draft. 
    AP, JA, EL, MW and PR, performed interpretation of the results, and reviewed and edited the draft. 
    All authors read and approved the final manuscript.

acknowledgements: |
    We thank the internal reviewers Bethan White and Xiaobing Zhou for their comments and feedback. 
    This work benefited from earlier unpublished work by Laura Davies, Phil Reid, Andrew G. Marshall. 
    This research was undertaken with the assistance of resources from the National Computational Infrastructure (NCI Australia), an NCRIS enabled capability supported by the Australian Government.
    This work was supported by ARC SRIEAS Grant SR200100005 Securing Antarctica’s Environmental Future.
# appendix: |
#   \section{Figures and tables in appendices}
#   Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
#   \subsection{Option 1}
#   If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
#   They will be correctly named automatically.
#   \subsection{Option 2}
#   If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.

@@

Accurate Antarctic sea-ice forecasts are crucial for climate monitoring and operational planning, yet they remain challenging due to model biases and complex ice-ocean-atmosphere interactions. 
The two versions of the Australian Bureau of Meteorology's ACCESS seasonal forecast system, +S1 and +S2, use identical model configuration and differ only in their initial conditions; primarily in that +S2 does not assimilate sea-ice observations, whereas +S1 does.   
This provides a convenient opportunistic experiment to assess the role of initial conditions on Antarctic sea-ice forecasts using more than 20 years of fully coupled simulations with two 9-member ensembles.
Our analysis reveals that both systems experience an extended melt season and delayed growth phase compared with observations. 
This leads to a significant negative sea-ice extent bias, which is corrected only in +S1 by the data assimilation system. 
The impact of the differing initial conditions on forecast errors varies dramatically by season: summer and autumn initial conditions (January-April) provide predictive skill for up to three months, with February initial conditions being particularly crucial. 
In contrast, winter forecasts of the two systems are statistically indistinguishable after only two weeks. 
Regional analysis of forecast skill suggests that this winter predictability barrier is most dramatic over East Antarctica, where even +S1 shows negative skill. 
These findings highlight the critical importance of comprehensive year-round sampling in predictability studies and suggest that operational sea-ice data assimilation efforts should prioritize the summer-autumn period when initial conditions have maximum impact on forecast skill.

>>>>

bibliography: references.bib  
running:
  title: The Importance of Initial Conditions in Seasonal Predictions of Antarctic Sea Ice
  author: Campitelli et al.
# This section is mandatory even if you declare that no competing interests are present.
competinginterests: |
  # Competing interests
  The authors declare no competing interests.
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
# copyrightstatement: |
#   The author's copyright for this publication is transferred to institution/company. 
### The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
### It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.
### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
availability:
  #code: |
  #  use this to add a statement when having only software code available
  #data: |
  #  use this to add a statement when having only data sets available
  codedata: |
    # Code/Data availability 
    The underlying code for this study is available on GitHub: https://github.com/eliocamp/access-s2_ice-eval.
    Raw data of +S1 and +S2 hindcast are not available due to size. 
    Derived datasets required to reproduce the results, including extent timeseries and error measures, are available in this Zenodo repository: https://zenodo.org/records/17479538 [@campitelli2025]
authorcontribution: |
    # Author contributions
    EC performed the data analysis and wrote the manuscript draft. 
    AP, JA, EL, MW and PR, performed interpretation of the results, and reviewed and edited the draft. 
    All authors read and approved the final manuscript.

acknowledgements: |
    # Acknowledgements
    We thank the internal reviewers Bethan White and Xiaobing Zhou for their comments and feedback. 
    This work benefited from earlier unpublished work by Laura Davies, Phil Reid, Andrew G. Marshall. 
    This research was undertaken with the assistance of resources from the National Computational Infrastructure (NCI Australia), an NCRIS enabled capability supported by the Australian Government.
    This work was supported by ARC SRIEAS Grant SR200100005 Securing Antarctica’s Environmental Future.
# appendix: |
#   \section{Figures and tables in appendices}
#   Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
#   \subsection{Option 1}
#   If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
#   They will be correctly named automatically.
#   \subsection{Option 2}
#   If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.

@@

Accurate Antarctic sea-ice forecasts are crucial for climate monitoring and operational planning, yet they remain challenging due to model biases and complex ice-ocean-atmosphere interactions. 
The two versions of the Australian Bureau of Meteorology's ACCESS seasonal forecast system, +S1 and +S2, use identical model configuration and differ only in their initial conditions; primarily in that +S2 does not assimilate sea-ice observations, whereas +S1 does.   
This provides a convenient opportunistic experiment to assess the role of initial conditions on Antarctic sea-ice forecasts using more than 20 years of fully coupled simulations with two 9-member ensembles.
Our analysis reveals that both systems experience an extended melt season and delayed growth phase compared with observations. 
This leads to a significant negative sea-ice extent bias, which is corrected only in +S1 by the data assimilation system. 
The impact of the differing initial conditions on forecast errors varies dramatically by season: summer and autumn initial conditions (January-April) provide predictive skill for up to three months, with February initial conditions being particularly crucial. 
In contrast, winter forecasts of the two systems are statistically indistinguishable after only two weeks. 
Regional analysis of forecast skill suggests that this winter predictability barrier is most dramatic over East Antarctica, where even +S1 shows negative skill. 
These findings highlight the critical importance of comprehensive year-round sampling in predictability studies and suggest that operational sea-ice data assimilation efforts should prioritise the summer-autumn period when initial conditions have maximum impact on forecast skill.

### 2025-12-11 analysis/paper/access-ice.tex
Adds section title to competing interests and etc...

\firstpage{1}
\maketitle
\begin{abstract}
Accurate Antarctic sea-ice forecasts are crucial for climate monitoring
and operational planning, yet they remain challenging due to model
biases and complex ice-ocean-atmosphere interactions. The two versions
of the Australian Bureau of Meteorology's ACCESS seasonal forecast
system, ACCESS-S1 and ACCESS-S2, use identical model configuration and
differ only in their initial conditions; primarily in that ACCESS-S2
does not assimilate sea-ice observations, whereas ACCESS-S1 does.\\
This provides a convenient opportunistic experiment to assess the role
of initial conditions on Antarctic sea-ice forecasts using more than 20
years of fully coupled simulations with two 9-member ensembles. Our
analysis reveals that both systems experience an extended melt season
and delayed growth phase compared with observations. This leads to a
significant negative sea-ice extent bias, which is corrected only in
ACCESS-S1 by the data assimilation system. The impact of the differing
initial conditions on forecast errors varies dramatically by season:
summer and autumn initial conditions (January-April) provide predictive
skill for up to three months, with February initial conditions being
particularly crucial. In contrast, winter forecasts of the two systems
are statistically indistinguishable after only two weeks. Regional
analysis of forecast skill suggests that this winter predictability
barrier is most dramatic over East Antarctica, where even ACCESS-S1
shows negative skill. These findings highlight the critical importance
of comprehensive year-round sampling in predictability studies and
suggest that operational sea-ice data assimilation efforts should
prioritize the summer-autumn period when initial conditions have maximum
impact on forecast skill.
\end{abstract}
\copyrightstatement{The author's copyright for this publication is
transferred to institution/company.}

@@

\codedataavailability{The underlying code for this study is available on
GitHub: https://github.com/eliocamp/access-s2\_ice-eval. Raw data of +S1
and +S2 hindcast are not available due to size. Derived datasets
required to reproduce the results, including extent timeseries and error
measures, are available in this Zenodo repository:
https://zenodo.org/records/17479538
\citep{campitelli2025}} %% use this section when having data sets and software code available

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontribution{EC performed the data analysis and wrote the
manuscript draft. AP, JA, EL, MW and PR, performed interpretation of the
results, and reviewed and edited the draft. All authors read and
approved the final manuscript.} %% optional section

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\competinginterests{The authors declare no competing
interests.} %% this section is mandatory even if you declare that no competing interests are present

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgements}
We thank the internal reviewers Bethan White and Xiaobing Zhou for their
comments and feedback. This work benefited from earlier unpublished work
by Laura Davies, Phil Reid, Andrew G. Marshall. This research was
undertaken with the assistance of resources from the National
Computational Infrastructure (NCI Australia), an NCRIS enabled
capability supported by the Australian Government. This work was
supported by ARC SRIEAS Grant SR200100005 Securing Antarctica's
Environmental Future.
\end{acknowledgements}

>>>>

\firstpage{1}
\maketitle
\begin{abstract}
Accurate Antarctic sea-ice forecasts are crucial for climate monitoring
and operational planning, yet they remain challenging due to model
biases and complex ice-ocean-atmosphere interactions. The two versions
of the Australian Bureau of Meteorology's ACCESS seasonal forecast
system, ACCESS-S1 and ACCESS-S2, use identical model configuration and
differ only in their initial conditions; primarily in that ACCESS-S2
does not assimilate sea-ice observations, whereas ACCESS-S1 does.\\
This provides a convenient opportunistic experiment to assess the role
of initial conditions on Antarctic sea-ice forecasts using more than 20
years of fully coupled simulations with two 9-member ensembles. Our
analysis reveals that both systems experience an extended melt season
and delayed growth phase compared with observations. This leads to a
significant negative sea-ice extent bias, which is corrected only in
ACCESS-S1 by the data assimilation system. The impact of the differing
initial conditions on forecast errors varies dramatically by season:
summer and autumn initial conditions (January-April) provide predictive
skill for up to three months, with February initial conditions being
particularly crucial. In contrast, winter forecasts of the two systems
are statistically indistinguishable after only two weeks. Regional
analysis of forecast skill suggests that this winter predictability
barrier is most dramatic over East Antarctica, where even ACCESS-S1
shows negative skill. These findings highlight the critical importance
of comprehensive year-round sampling in predictability studies and
suggest that operational sea-ice data assimilation efforts should
prioritise the summer-autumn period when initial conditions have maximum
impact on forecast skill.
\end{abstract}

@@

\codedataavailability{\section{Code/Data
availability}\label{codedata-availability}

The underlying code for this study is available on GitHub:
https://github.com/eliocamp/access-s2\_ice-eval. Raw data of +S1 and +S2
hindcast are not available due to size. Derived datasets required to
reproduce the results, including extent timeseries and error measures,
are available in this Zenodo repository:
https://zenodo.org/records/17479538
\citep{campitelli2025}} %% use this section when having data sets and software code available

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontribution{\section{Author
contributions}\label{author-contributions}

EC performed the data analysis and wrote the manuscript draft. AP, JA,
EL, MW and PR, performed interpretation of the results, and reviewed and
edited the draft. All authors read and approved the final
manuscript.} %% optional section

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\competinginterests{\section{Competing
interests}\label{competing-interests}

The authors declare no competing
interests.} %% this section is mandatory even if you declare that no competing interests are present

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgements}
\section{Acknowledgements}\label{acknowledgements}

We thank the internal reviewers Bethan White and Xiaobing Zhou for their
comments and feedback. This work benefited from earlier unpublished work
by Laura Davies, Phil Reid, Andrew G. Marshall. This research was
undertaken with the assistance of resources from the National
Computational Infrastructure (NCI Australia), an NCRIS enabled
capability supported by the Australian Government. This work was
supported by ARC SRIEAS Grant SR200100005 Securing Antarctica's
Environmental Future.
\end{acknowledgements}
